---
title: "Data Visualization"
author: "Curtis Donovan"
date: "8/21/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Visualization

### 3.1 Introduction

```{r}
# we must load the tidyverse library every session
library(tidyverse)
```

### 3.2 First Steps

Question: Do cars with big engines use more gas then cars with small engines?

##### 3.2.1 The mpg data frame

```{r}
mpg
```

### 3.2.2 creating a ggplot

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy))

```

##### Exercises

1. Run ggplot(data = mpg). What do you see?

```{r}
ggplot(data = mpg)
```
A grey rectangle, empty canvas.

2. How many rows are in mpg? How many columns?

```{r}
?mpg
```

234 rows and 11 columns

3. What does the drv variable describe? Read the help for ?mpg to find out.

```{r}
?mpg
```

The drv variable desribes the type of drive that each car can be, front-wheel, four-wheel, or rear-wheel.

4. Make a scatterplot of hwy vs cyl.

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = hwy, y = cyl))
```

5. What happens if you make a scatterplot of class vs drv? Why is the plot not useful?

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = class, y = drv))
#Because it is just showing you the data again
```
Both variables are categorical. 


### 3.3 Aestetic Mappings

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = class))
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, size = class))
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, alpha = class))
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, shape = class))
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

### 3.3.1 Exercises

1. What’s gone wrong with this code? Why are the points not blue?

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x =displ, y = hwy, color = "blue"))

```

Because the blue is supposed to be outside of the parentheses in this case to make the points blue.

2.Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg?

```{r}
?mpg

```

Categorical are the non numerical variables with a limit of items in the variable. You can see which values are numerical.

3. Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x =displ, y = hwy, color = cty))
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, size = cty))
```


The shape aesthetic can't handle continuous variables because continious variables need a much broader range.

4. What happens if you map the same variable to multiple aesthetics?

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x =displ, y = hwy, color = class, shape = class))
```

Both aesthetics are done to the same variable. 

6. What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)? Note, you’ll also need to specify x and y.

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x =displ, y = hwy, colour = displ < 5))

```
It makes a true vs false comparison.

### 3.5 Facets

```{r}
ggplot(data = mpg) + geom_point(aes(x = displ, y = hwy)) + facet_wrap(~ class, nrow = 2)
```

```{r}
ggplot(data = mpg) + geom_point(aes(x = displ, y = hwy)) + facet_grid(drv ~ cyl)
```

### 3.5.1 Exercises

1.What happens if you facet on a continuous variable?

```{r}
ggplot(data = mpg) + geom_point(aes(x = displ, y = hwy)) + facet_wrap(~ cty, nrow = 2)

```

It attempts to make the variable categorical.

2. What do the empty cells in plot with facet_grid(drv ~ cyl) mean? How do they relate to this plot?

```{r}
ggplot(data = mpg) + geom_point(aes(x = displ, y = hwy)) + facet_grid(drv ~ cyl)

```
That means at that amount of cylinders there is no cars of those types of drives


3. What plots does the following code make? What does . do?

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) +facet_grid(drv ~ .)

```
Ignores dimension.

4. Take the first faceted plot in this section: 
What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_wrap(~ class, nrow = 2)

```
#It lets you analyze a specific variable's data set better, but it also doesn't let you see the data compared to one another as well.

5. Read ?facet_wrap. What does nrow do? What does ncol do? What other options control the layout of the individual panels? Why doesn’t facet_grid() have nrow and ncol arguments?

```{r}
?facet_wrap
?facet_grid
```
Nrow and ncol choose number of rows and columns. Facet-grid() doesn't have nrow and ncol because the variables choose the columns and rows.

6. When using facet_grid() you should usually put the variable with more unique levels in the columns. Why?

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_grid(cyl ~ cty)
```

Because there is more space.


### 3.6 Geometric Objects

```{r}
ggplot(data = mpg) + geom_smooth(mapping = aes(x = displ, y = hwy))
```

```{r}
ggplot(data = mpg) + geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))
```

```{r}
ggplot(data = mpg) + geom_smooth(mapping = aes( x = displ, y = hwy, group = drv))
```

```{r}
ggplot(data = mpg) + geom_smooth(mapping = aes( x = displ, y = hwy, color = drv))
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + geom_smooth(mapping = aes(x = displ, y = hwy))
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = drv)) + geom_smooth(mapping = aes(x = displ, y = hwy, color = drv))
```
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + geom_point() + geom_smooth()
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + geom_smooth() + geom_point(mapping = aes(color = class))
```

### 3.6.1 Exercises 

1. What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?
Line chart:geom_line()
Boxplot:geom_boxplot() 
Histogram: geom_histogram()
Area Chart: geom_area()
```{r}
??geom
```

2. Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.

Graph that shows the displ compared to the hwy with drv disinguishing showing line of best fit through each drive type.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```


3. What does show.legend = FALSE do? What happens if you remove it?
Why do you think I used it earlier in the chapter?

The legend wouldn't show up.

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = drv, show.legend = FALSE))
```

4. What does the se argument to geom_smooth() do?

It removes the extra area that it could be from the smooth.

5. Will these graphs look different? Why/why not?

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()
```

```{r}
ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

No because the mappings and aesthetics are the same.


### 3.7 Statistical Transformations

```{r}
?diamonds
```
```{r}
ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut))
```
### 3.8 Position Adjustments
```{r}
ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut, color = cut))
```

```{r}
ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut, fill = cut))
```

```{r}
ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut, fill = clarity))
```

```{r}
ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) + geom_bar(alpha = 0.2, position = "identity")
```

```{r}
ggplot(data = diamonds, mapping = aes(x = cut, color = clarity)) + geom_bar(fill = NA, position = "identity")
```

```{r}
ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) + geom_bar(position = "fill")
```

```{r}
ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) + geom_bar(position = "dodge")
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy))
```


```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy), position = "jitter")
```


### 3.8.1 Exercises

Notebook check - exercises & notes

Project - start next week / due week after

1. What is the problem with this plot? How could you improve it?
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + geom_point()
```
Ans: The problem with this plot is you can't see all the points, you could improve it by using jitter.

2. What parameters to geom_jitter() control the amount of jittering?

Ans: The amount of width and height to the jitter control the jittering.

3. Compare and contrast geom_jitter() with geom_count().

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + geom_jitter()
```
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + geom_count()
```

Ans: Both graphs are very cluttery, but geom_count shows the amount of points using size and geom_jitter tries to show all the points.

4. What’s the default position adjustment for geom_boxplot()? Create a visualisation of the mpg dataset that demonstrates it.

```{r}
ggplot(data = mpg) + geom_boxplot(mapping = aes(x = class, y = hwy))
?geom_boxplot
```
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = class, y = hwy))
```

Ans: It accompanies as many points as it can within a certain range.

