---
title: "Data Wrangling"
author: "Curtis Donovan"
date: "9/18/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 5.1 Introduction

### 5.1.1 Prerequisites

```{r}
library(tidyverse)
library(nycflights13)
library(Lahman)
```

### 5.1.2 nycflights13

```{r}
flights
```
### 5.1.3 dplyr

verbs

filter
arrange
select
mutate
summarize

verb(database, whatTodo?)

# 5.2 Filter rows with filter()

Flights that occured on Jan 1st

```{r}
filter(flights, month == 1, day == 1)
```

```{r}
jan1 <- filter(flights, month == 1, day == 1)

```


```{r}
(Dec25 <- filter(flights, month == 12, day == 25))
```

### 5.2.1 Comparisons

==, >, >=, < <=, !

```{r}
filter(flights, month == 1)
```


```{r}
cos(pi/2) == 0
```

```{r}
near(cos(pi/2), 0)
```

### 5.2.2 Locical Operators

or -------> |
and -------> &
not --------> !

Flights that departed in November **or** December

```{r}
filter(flights, month == 11 | month == 12)
```

May, June, July, August

```{r}
filter(flights, month == 5 | month == 6 | month == 7 | month == 8)
```

```{r}
filter(flights, month %in% c(5, 6, 7, 8))
```

Flights that departed in November and arrive in December

```{r}
#tbd
```

Find the flights that weren't delayed (on departure or arrival) by more than two hours

```{r}
filter(flights, dep_delay <= 120 & arr_delay <= 120)
```

##### 5.2.3 Missing Values

```{r}
100 == NA
2 > NA
x <- NA
is.na(x)
```


How do I get rid of NA's on my dataset?

```{r}
filter(flights, is.na(dep_time))
filter(flights, !is.na(dep_time))
withoutNA <- filter(flights, !is.na(dep_time))
withoutNA

filter(withoutNA, is.na(dep_time)) 
```

##### 5.2.4 Exercises

1. Find all flights that had an arrival delay of 2 or more hours.

```{r}
filter(flights, arr_delay >= 120)
```

2. All flights that flew to Houston.

```{r}
filter(flights, dest == "IAH" | dest == "HOU")
filter(flights, dest %in% c("IAH", "HOU"))
```

3. Opperated by United, American, or Delta. DL, AA, UA.

```{r}
filter(flights, carrier == "AA" | carrier == "DL" | carrier == "UA")
```

4. Departed in Summer.

```{r}
filter(flights, month == 7 | month == 8 | month == 9)
```

5. Arrived more than 2 hours late.

```{r}
filter(flights, arr_delay > 120 & dep_delay <= 0)
filter(flights, arr_delay > 120, dep_delay <= 0)
```

Were delayed by at least an hour, but made up over 30 minutes in the flight.

```{r}
filter(flights, dep_delay >= 60, dep_delay - arr_delay > 30)
```

2. Departed between midnight and 6am using variable between().

```{r}
filter(flights, between(dep_time, 00, 600))
```
3. How many flights have a missing dep_time? What other variables are missing? What might these rows represent?

```{r}
filter(flights, is.na(dep_time))
```
Also missing the departure delay and the arrival time and arrival delay.

Arrange Rows using arrange()

```{r}
arrange(flights, dep_delay)
arrange(flights, dep_delay, month)

```

Use desc() to re-order by a column in descending order
```{r}
arrange(flights, desc(dep_delay))
```

To get rid of NAs
```{r}
newDataBase <- filter(flights, !is.na(dep_delay))
```

Shortest flights

```{r}
arrange(flights, air_time)
```

##### 5.3.1 Exercises

Sort to find fastest flights (highest speed)
```{r}
arrange(flights, desc(distance / air_time))
```

Which flights travelled the farthest? Which travelled the shortest?

```{r}
arrange(flights, desc(distance))
arrange(flights, distance)
```


To isolate a variable use select()
```{r}
select(flights, year, month, day)
select(flights, year:day)
#To isolate all variables besides day year and month use -()
select(flights, -(year:day))
```

Select() has a lot of helper functions you can use within.

starts_with()
ends_with()
containts()

```{r}
select(flights, starts_with('dep'))
select(flights, ends_with('lay'))
select(flights, contains('_'))
```

There is also a rename function that can be used to rename data sets and variables. rename()

```{r}
#rename(database, new_name = old_name)
rename(flights, tail_num = tailnum)
```

Select() also has a helper function everything().

```{r}
select(flights, dep_delay, arr_delay, everything())
```


##### 5.4.1 Exercises

1. Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay.

```{r}
select(flights, starts_with('dep') | starts_with('arr'))
```

2. What happens if you include the name of a variable multiple times in a select() call?

```{r}
select(flights, air_time, distance, air_time)
```


### 5.5 Add new variables with mutate()

```{r}
flights_narrow <- select(flights, year:day, ends_with('delay'), distance, air_time)
```

```{r}
flights_narrow
```

```{r}
mutate(flights_narrow, gain = dep_delay - arr_delay, speed = distance / air_time * 60)
```

```{r}
mutate(flights_narrow, gain = arr_delay - dep_delay, speed = distance / air_time * 60, gain_per_hour = gain / air_time)
```


The transmute variable isolates the new columns you create.
```{r}
transmute(flights, gain = dep_delay - arr_delay, speed = distance / air_time * 60, gain_per_hour = gain / air_time)
```

### 5.6 Grouped summaries with summarize()

Collapses all data to a single row

```{r}
summarize(flights, delay = mean(dep_delay, na.rm = TRUE))
```

Summarize() isn't super useful unless you combine it with group_by().

```{r}
by_day <- group_by(flights, year, month, day)

summarize(by_day, delay = mean(dep_delay, na.rm = TRUE))
```

```{r}
by_carrier <- group_by(flights, carrier)

summarize(by_carrier, delay = mean(dep_delay, na.rm = TRUE))
```

```{r}
by_airTime <- group_by(flights, air_time)

summarize(by_airTime, delay = mean(dep_delay, na.rm = TRUE))
```

```{r}
flights
by_airport <- group_by(flights, origin)
summarize(by_airport, delay = mean(dep_delay, na.rm = TRUE))
```

##### 5.6.1 Combining multiple operations with the pipe

Group flights by destination

```{r}
by_dest <- group_by(flights, dest)
```


Summarize the group, using average distance, average delay, and count:

```{r}
delays <- summarize(by_dest, count = n(), 
dist = mean(distance, na.rm = TRUE),
delay = mean(arr_delay, na.rm = TRUE))

delays
```

Make a scatter plot of distance vs delay, with count mapped to the size aesthetic.

```{r}
ggplot(data = delays) + geom_point(mapping = aes(x = dist, y = delay, size = count))
```


```{r}
delays <- filter(delays, count > 20, dest != 'HNL')
```


Clean up your plot:

```{r}
ggplot(data = delays, mapping = aes(x = dist, y = delay)) + 
  geom_point(aes(size = count), alpha = .3) +
  geom_smooth(se = FALSE) #se = FALSE removes standard error band
```

The longer the flight, the less delay.

```{r}
by_dest <- flights %>%
  group_by(dest) %>%
  summarize(count =n(), dist = mean(distance, na.rm = TRUE),
delay = mean(arr_delay, na.rm = TRUE)) %>%
filter(count > 20, dest != 'HNL')

ggplot(data = delays, mapping = aes(x = dist, y = delay)) + 
  geom_point(aes(size = count), alpha = .3) +
  geom_smooth(se = FALSE) #se = FALSE removes standard error band
```

What does na.rm mean, remove the NA what happens if we don't set it?

```{r}
flights %>%
  group_by(year, month, day) %>%
  summarize(delay = mean(dep_delay))

flights %>%
  group_by(year, month, day) %>%
  summarize(delay = mean(dep_delay, na.rm = TRUE))
```

is.na(variable) - asks a question - is the variable NA?

na.rm = TRUE - gives a command - remove all NA

```{r}
not_cancelled <- flights %>%
  filter(!is.na(arr_delay) & !is.na(dep_delay))

not_cancelled
```

##### 5.6.3 Counts

```{r}
delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarize(delay = mean(arr_delay))

delays

arrange(delays, desc(delay))
```

```{r}
delays <- not_cancelled %>%
  group_by(tailnum) %>%
  summarize(
    delay = mean(arr_delay),
    count = n()
  )

delays

arrange(delays, desc(delay))

ggplot(data = delays) + 
  geom_point(mapping = aes(x = count, y = delay), alpha = 1/3)
```

```{r}
delays %>%
  filter(count > 25) %>%
  ggplot() + 
  geom_point(mapping = aes(x = count, y = delay), alpha = .1)
```

```{r}
Batting
```

```{r}
batting <- Batting

batters <- batting %>%
  group_by(playerID) %>%
  summarize(
   ba = sum(H, na.rm = TRUE) / sum(AB, na.rm = TRUE),
   ab = sum(AB)
  )

batters

arrange(batters, desc(ba))

batters %>%
  filter(ab > 100) %>%
  ggplot(mapping = aes(x = ab, y = ba)) +
  geom_point() + 
  geom_smooth(se = FALSE)
```










